\documentclass{llncs}

\usepackage[american]{babel}
\usepackage{graphicx}
\usepackage[T1]{fontenc}

\usepackage[
bookmarks=false,
breaklinks=true,
colorlinks=true,
linkcolor=black,
citecolor=black,
urlcolor=black,
pdfpagelayout=SinglePage
]{hyperref}
% enables correct jumping to figures when referencing
% \usepackage[all]{hypcap}

% to typeset URLs, URIs, and DOIs
\usepackage{url}
\def\UrlFont{\rmfamily}
\usepackage{epstopdf}
\usepackage[]{subfig}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{amssymb}

\usepackage{bbding}


\begin{document}

\title{Efficient Algebraic Multigrid Preconditioners \\on Clusters of GPUs}
%
%\titlerunning{Hamiltonian Mechanics}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Ambra Abdullahi Hassan\inst{1}%
\and Valeria Cardellini\inst{1} \and Pasqua D'Ambra\inst{2} \and \\ 
Daniela di Serafino\inst{3} \and  Salvatore Filippone\inst{4}}
%
%\authorrunning{} % abbreviated author list (for running head)
%
\institute{Universit\`a degli Studi di Roma ``Tor Vergata'', Roma, Italy\\
 \email{\{ambra.abdullahi,cardellini\}@uniroma2.it}
\and
Istituto per le Applicazioni del Calcolo ``Mauro Picone'', CNR, Napoli, Italy\\
 \email{pasqua.dambra@cnr.it}
\and
Universit\`a degli Studi della Campania ``Luigi Vanvitelli'', Caserta, Italy
\email{daniela.diserafino@unicampania.it}
\and
 Cranfield University, Cranfield, UK\\
 \email{salvatore.filippone@cranfield.ac.uk}
}
\maketitle              % typeset the title of the contribution

\begin{abstract}
Many scientific applications require the solution of large and sparse linear systems of equations using Krylov
subspace methods; in these cases, the choice of a good preconditioner may be crucial for the convergence
of the Krylov solver. Algebraic multigrid methods are widely used as preconditioners, because of their optimal
computational cost and their algorithmic scalability. The wide availability of GPUs, now found in many
of the fastest supercomputers, poses the problem of implementing efficiently these algorithms on
high-throughput processors. In this work we focus on the choice and implementation of smoothers capable
of exploiting the computational power of clusters of GPUs. We consider block-Jacobi methods using sparse
approximate inverses in the solve phase concerning the local blocks. The choice of approximate inverses 
instead of sparse matrix factorizations is driven by the large amount of parallelism exposed by the matrix-vector
product as compared to the solution of large triangular systems on GPUs. The selected smoother and solvers are
implemented within the algebraic multigrid preconditioning framework provided by the MLD2P4 library, using sparse
matrix data structures and Krylov solvers from the PSBLAS library. Their behaviour is illustrated
in terms of execution speed and exploitation of computational cores, as well as of algorithmic scalability,
on a test case concerning groundwater modelling, provided by J\"ulich Supercomputing Center within the
Horizon 2020 Project EoCoE.
\end{abstract}
%
\input{1-intro.tex}
%
\input{2-amg-prec.tex}
%
\input{3-gpu-linalg.tex}
%
\input{4-experiments.tex}
%
\input{5-concl.tex}
%
% ---- Bibliography ----
%
\bibliographystyle{splncs03}
\bibliography{europar,ainv}

% \begin{thebibliography}{6}
% %

% \bibitem {}


% \end{thebibliography}
\end{document}
